### 2024-10-11 draft-2/draft-2.tex
added the position message

@@

\minor The engine's standard input and output must be open file descriptors and
the engine's standard error must be an open file descriptor until closed by the
engine. The sequence of bytes that the client sends to the engine via the
engine's standard input is the \term{client stream}. The sequence of bytes
that the engine sends to the client via the engine's standard output is the
\term{engine stream}.
\vskip-\parskip
\greybox{There are no requirements for an engine's standard error except that it
be open. For example, the engine's standard error may be directed to a null
file, to the client's standard output or standard error, or to a log file. There
are accordingly no restrictions on the sequence of bytes that the engine writes
to its standard error.}
\vskip-\parskip
\bluebox{No particular encoding is specified: the client and engine streams are
not required to be encoded sequences of Unicode scalar values (or code points
of any other character set). The specification instead governs the client and
engine streams as sequences of bytes {\it per se}.\par
However, clients and engines are recommended to use {\caps utf-8} for
client−engine communication so that engine, author, and option names are
displayed properly. Note that clients and engines may use different encodings
for other interfaces, such as for file system paths that are passed as arguments
to the operating system. For example, the client and engine may be communicating
over a network connexion and the client may be running on Linux (where paths
are arbitrary byte sequences that do not contain \hex{00} or \hex{2f}) but the
engine may be running on Windows (where {\caps ntfs} is a common filesystem,
which stores file names in {\caps utf-16}, but where older libraries or
{\caps api}s may expect paths encoded in the local code page), and so the
path of a tablebase file may require transcoding, implicitly or explicitly.}

@@

\vskip-\parskip
\greybox{A valid position (as defined above) need not be reachable from the
starting position.\par
Some engines designed for gameplay rather than analysis may additionally require
that the positions they are sent must be reachable from the starting position.
Such constraints are specific to engines, and are not constraints of the
Universal Chess Interface.}
\vskip-\parskip
\greybox{The 2023 {\caps fide} Laws of Chess state that the game ends
immediately when a player cannot checkmate the king by any series of legal
moves, but this condition does not cause a position to not be valid (as defined
above).}
\vskip-\parskip
\bluebox{Engines are recommended to accept positions that are checkmate or
stalemate even though they are not required to handle receiving such positions.
For such positions, engines should report the null move as the best move (see
\the\majorcount·{\advance\minorcount by 7 \the\minorcount} below).}

\minor A \term{move} is either a tuple with two fields – a board index
called the \term{source} and a board index called the \term{destination}
– or a tuple with three fields: a~source, a~destination, and a kind called
the \term{promotion kind} that is either queen, rook, bishop, or knight.

@@

\minor A \term{Forsyth–Edwards Notation ({\caps fen}) record} is a sequence of
six tokens ⟨{\it board, side to move, rights, {\icap ep} target, {\icap dfz},
move number}⟩ of the form\par
%
\dimen0=5em
\vskip-0.5\parskip\hskip3ex\rjust{\dimen0}{\it board}
  \dt=\dt\ \r{(['{\it row}']*8).join('\pt\s/\pt')}\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it row}
  \dt=\dt\ \s8 \alt\ [\s1–\s7\s{KQRBNPkqrbnp}]+\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it side to move}
  \dt=\dt\ \s w \alt\ \s b\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it rights}
  \dt=\dt\ \s- \alt\ \s K?\pt\s Q?\pt\s k?\pt\s q?\par
  % \lower0.5pt\hbox{\notomono ∩} [\s{KQkq}]+)\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it {\icap ep} target}
  \dt=\dt\ \s- \alt\ [\s a–\s h][\s{36}]\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\icap dfz}
  \dt=\dt\ \s0 \alt\ [\s1–\s9][\s0–\s9]\lower0.25ex\hbox{*}\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it move number}
  \dt=\dt\ [\s1–\s9][\s0–\s9]\lower0.25ex\hbox{*}\par
\vskip-\parskip\hskip3ex
  \comment{The dash “\s-” is \hex{2d}.}

@@

\vskip-0.5\parskip The bytes \s K and \s k must each occur exactly once in
{\it board}.

\vskip-0.5\parskip The byte \s P must not occur in the first {\it row} and must
not occur in the last {\it row}, and likewise for the byte \s p.

\vskip-0.5\parskip The {\icap dfz} token interpreted as a decimal integer must
be 100 or less.

\vskip-0.5\parskip The {\it move number} token interpreted as a decimal integer
must be less than 10\pt000.\par}

@@

\minor A \term{move token} is a token of the form
\hbox{[\s a–\s h][\s1–\s8][\s a–\s h][\s1–\s8]} or\hfil\penalty-10000
\hbox{[\s a–\s h]\s7[\s a–\s h]\s8[\s{qrbn}]} or
\hbox{[\s a–\s h]\s2[\s a–\s h]\s1[\s{qrbn}]}.

\minor A \term{null move token} is a token of the form \s{0000}.

@@

\minor A \s{uci}, \s{ucinewgame}, \s{isready}, \s{stop}, or \s{quit} message
is well-formed if it contains a single token.

\minor A \s{debug} message is well-formed if it contains two tokens,
either ⟨\s{debug on}⟩ or ⟨\s{debug off}⟩.
\vskip-\parskip
\bluebox{If engines have debugging information to expose via the client
(that is, through standard output rather than standard error), they should
feely send \s{info string} messages with this information after receiving
\s{debug on} and suppress these messages after receiving \s{debug off}.}

\minor A \s{setoption} message is well-formed if it has a prefix
of the form ⟨\s{setoption name} \ellipsis\ \s{value}⟩, where the ellipsis is
one or more tokens (none of which are \s{value}) matching an option name listed
in an \s{option} message enqueued by the engine, and a suffix satisfying the
condition(s) for the corresponding option type:

@@

\vskip-0.5\parskip If \s{check}, the suffix contains a single token, either
\s{true} or \s{false}.

@@

\vskip-0.5\parskip A \s{setoption} message is also well-formed if it is of the
form ⟨\s{setoption name} \ellipsis\pt⟩ where the ellipsis is one or more tokens
(none of which are \s{value}) matching an option name listed in an \s{option}
message enqueued by the engine for which the corresponding option type is
\s{button}.

\minor A \s{position} message is well-formed if\ellipsis

>>>>

\def\subfamily{Alegreya Sans Medium}%
\def\subsize{8}%
\font\subrm=\ldfont {\subfamily}{\subsize}{\basefeat}%
\font\subit=\ldifont{\subfamily}{\subsize}{\basefeat}%
\def\sub#1{\lower 0.25ex\hbox{\subrm #1}}%
\def\isub#1{\lower 0.25ex\hbox{\subit #1}}%

@@

\minor The engine's standard input and output must be open file descriptors and
the engine's standard error must be an open file descriptor until closed by the
engine. The sequence of bytes that the client sends to the engine via the
engine's standard input is the \term{client stream}. The sequence of bytes
that the engine sends to the client via the engine's standard output is the
\term{engine stream}.
\vskip-\parskip
\greybox{There are no requirements for an engine's standard error except that
it be open. For example, the engine's standard error may be directed to a null
device, to the client's standard output or standard error, or to a log file.
There are accordingly no restrictions on the sequence of bytes that the engine
writes to its standard error.}
\vskip-\parskip
\bluebox{No particular encoding is specified: the client and engine streams are
not required to be encoded sequences of Unicode scalar values (or code points
of any other character set). The specification instead governs the client and
engine streams as sequences of bytes {\it per se}.\par
However, clients and engines are recommended to use {\caps utf-8} for
client−engine communication so that engine, author, and option names are
displayed properly. Note that clients and engines may use different encodings
for other interfaces, such as for file system paths that are passed as arguments
to the operating system. For example, the client and engine may be communicating
over a network connexion and the client may be running on Linux (where paths
are arbitrary byte sequences that do not contain \hex{00} or \hex{2f}) but the
engine may be running on Windows (where {\caps ntfs} is a common filesystem,
which stores file names in {\caps utf-16}, but where older libraries or
{\caps api}s may expect paths encoded in the local code page), and so the
path of a tablebase file may require transcoding, implicitly or explicitly.}

@@

\vskip-\parskip
\greybox{A valid position (as defined above) need not be reachable from the
starting position.\par
Some engines designed for gameplay rather than analysis may additionally require
that the positions they are sent must be reachable from the starting position.
Such constraints are specific to engines, and are not constraints of the
Universal Chess Interface.}
\vskip-\parskip
\greybox{The 2023 {\caps fide} Laws of Chess state that the game ends
immediately when a player cannot checkmate the king by any series of legal
moves, but this condition does not cause a position to not be valid (as defined
above).}
\vskip-\parskip
\bluebox{Engines are recommended to accept positions that are checkmate or
stalemate even though they are not required to handle receiving such positions.
For such positions, engines should report null as the best move (see
\the\majorcount·{\advance\minorcount by 8 \the\minorcount} below).}

\minor A \term{move} is a tuple of three values: a board index called the
\term{source}, a board index called the \term{destination}, and a field
called the \term{promotion kind} that is either {\caps none} or queen, rook,
bishop, or knight.

@@

\minor A \term{Forsyth–Edwards Notation ({\caps fen}) record} is a sequence of
six tokens ⟨{\it board, side to move, rights, {\icap ep} target, {\icap dfz},
move number}⟩ of the form\par
%
\dimen0=5em
\vskip-0.5\parskip\hskip3ex\rjust{\dimen0}{\it board}
  \dt=\dt\ \r{(['{\it row}']*8).join('\pt\s/\pt')}\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it row}
  \dt=\dt\ \s8 \alt\ [\s1–\s7\s{KQRBNPkqrbnp}]+\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it side to move}
  \dt=\dt\ \s w \alt\ \s b\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it rights}
  \dt=\dt\ \s- \alt\ \s K?\pt\s Q?\pt\s k?\pt\s q?\par
  % \lower0.5pt\hbox{\notomono ∩} [\s{KQkq}]+)\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it {\icap ep} target}
  \dt=\dt\ \s- \alt\ [\s a–\s h][\s{36}]\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\icap dfz}
  \dt=\dt\ \s0 \alt\ [\s1–\s9][\s0–\s9]?\ \alt\ \s{100}\par
\vskip-\parskip\hskip3ex\rjust{\dimen0}{\it move number}
  \dt=\dt\ [\s1–\s9][\s0–\s9]\char"7B\ot0,\pt3\ot\char"7D\par
\vskip-\parskip\hskip3ex
  \comment{The dash “\s-” is \hex{2d}.}

@@

@@

\minor A \term{move token} is a token of the form\par
\vskip-\parskip\hskip3ex
  [\s a–\s h][\s1–\s8][\s a–\s h][\s1–\s8][\s{qrbn}]?\par
\vskip-\parskip The first and second bytes of a move token interpreted in the
natural way as a board index is the move token’s \term{source}, the third and
fourth bytes of a move token interpreted in the natural way as a board index is
the move token’s \term{destination}, and the fifth byte (if present) interpreted
in the natural way as a kind is the \term{promotion kind}. If the fifth byte is
not present, the promotion kind of the move token is {\caps none}.

\minor For a given position {\it P}, a move token {\it T} \term{denotes} a
move {\it M} if {\it M} is valid for {\it P} and the source, destination, and
promotion kind of {\it T} match the source, destination, and promotion kind of
{\it M} (respectively).

\minor A \term{null token} is a token of the form \s{0000}.

@@

\minor A \s{uci}, \s{ucinewgame}, \s{isready}, \s{stop}, or \s{quit} message
is well-formed if it contains exactly one token.

\minor A \s{debug} message is well-formed if it contains exactly two tokens,
either ⟨\ot\s{debug on}\ot⟩ or ⟨\ot\s{debug off}\ot⟩.
\vskip-\parskip
\bluebox{If engines have debugging information to expose via the client
(that is, through standard output rather than standard error), they should
feely send \s{info string} messages with this information after receiving
\s{debug on} and suppress these messages after receiving \s{debug off}.}

\minor A \s{setoption} message is well-formed if it has a prefix
of the form ⟨\ot\s{setoption name} {\it tok}+ \s{value}\ot⟩, where {\it tok}+ is
one or more tokens (none of which are \s{value}) matching an option name listed
in an \s{option} message enqueued by the engine, and a suffix satisfying the
condition(s) for the corresponding option type:

@@

\vskip-0.5\parskip If \s{check}, the suffix contains exactly one token, either
\s{true} or \s{false}.

@@

\vskip-0.5\parskip A \s{setoption} message is also well-formed if it is of the
form ⟨\ot\s{setoption name} {\it tok}+\pt⟩ where {\it tok}+ is one or more tokens
(none of which are \s{value}) matching an option name listed in an \s{option}
message enqueued by the engine for which the corresponding option type is
\s{button}.

\minor A \s{position} message is well-formed if it has the form\par
\vskip-\parskip\hskip3ex
⟨\pt\s{position}
  (\ot\s{startpos} \alt\ \s{fen} {\it fen}\ot⁶\pt)
  (\ot\s{moves} {\it mov}+\ot)?\pt\ot⟩\par
\vskip-\parskip where {\it fen}\ot⁶ is an {\caps fen} record and {\it mov}+ is
one or more move tokens. If~\s{fen} is specified, the position described by
{\it fen}\ot⁶ must be valid. If \s{moves} is specified:

\vskip-0.5\parskip Let {\it P}₀ be the starting position if \s{startpos} is
specified and the position described by {\it fen}\ot⁶ if \s{fen} is specified.
Let {\it N} be the number of move tokens specified.

\vskip-0.5\parskip For {\it n} from 0 to {\it N}−1 inclusive, the {\it n}-th
token {\it T}\isub{n} of {\it mov}+ must denote a move {\it M}\isub{n} from
{\it P}\isub{n}. Then let {\it P}\sub{{\subit n}+1} be the position that
immediately follows from {\it P}\isub{n} when {\it M}\isub{n} is applied.
